Reproducible Data Peer Assessment 1
=====================

First we load the data:

```{r echo = T}
data <- read.csv('activity.csv')
str(data)
```

Looks like read.csv pulled in the dates as factors -- let's convert them to
dates:

```{r echo = T}
data$date <- as.Date(data$date)
str(data)
```

Great.  Now, let's find out some info about total number of steps taken per day.

Let's look at mean total number of steps taken per day.  First we get the unique
dates, then we get the total steps for each day (ignoring na's).  Finally, we
make a histogram plot of the total counts.

```{r echo = T}
unique_dates <- unique(data$date)
date_total <- function (d) sum(na.exclude = T, data$steps[data$date == d])
daily_steps <- sapply(unique_dates, date_total)
hist(daily_steps)
```

Now let's summarize the daily_steps to see the mean and median:

```{r echo = T}
summary(daily_steps)
```

Now let's look at average steps taken by time interval.  First, like with days,
we'll get the unique time intervals.  Then we'll average over all days the
number of steps taken in each of those intervals (excluding na's). Finally, we
can plot the results:

```{r echo = T}
time_intervals <- unique(data$interval)
mean_by_itvl <- function (i) mean(na.rm = T, data$steps[data$interval == i])
mean_steps_by_itvl <- sapply(time_intervals, mean_by_itvl)
plot(time_intervals, mean_steps_by_itvl)
```

We can also find the time interval with the maximum average number of steps
across the days:

```{r echo = T}
time_intervals[mean_steps_by_itvl == max(mean_steps_by_itvl)]
```

Looks like it's 8:35-8:40 in the morning.

Let's think about how to deal with the missing values in our dataset.
First, how many are there?

```{r echo = T}
sum(is.na(data$steps))
nrow(data)
sum(is.na(data$steps)) / nrow(data)
```

Maybe those na's are biasing our findings.  Let's fill each na with the average
number of steps for that time interval over the available days, and recalculate
the mean and median number of steps per day:

```{r echo = T}
named_mean_steps_by_itvl <- setNames(mean_steps_by_itvl, 
                                     as.character(time_intervals))
lookup_mean <- function (i) {
        named_mean_steps_by_itvl[[(as.character(data$interval[[i]]))]] }
filled_steps <- sapply(1:length(data$steps), 
                       function (i) { 
                               if (!is.na(data$steps[[i]]))
                                       data$steps[[i]] 
                               else lookup_mean(i)})
filled_date_total <- function (d) sum(na.exclude = T, filled_steps[data$date == d])
filled_daily_steps <- sapply(unique_dates, filled_date_total)
hist(filled_daily_steps)
summary(filled_daily_steps)
```

This filling raises the 1st quartile and lowers the 3rd, but leaves the min,
max, median, and mean the same.  It slightly raises the average of the total
daily number of steps.

Finally, lets compare weekends and weekdays, using the filled steps data:

```{r echo = T}
data_w_filled_steps <- data
data_w_filled_steps$steps <- filled_steps
days_of_week <- sapply(data_w_filled_steps$date, function (d) weekdays(d))
is_weekend <- as.factor(sapply(days_of_week, function (d) d == 'Sunday' | d == 'Saturday'))
split_by_weekend <- split(data_w_filled_steps, is_weekend)
weekends_only <- split_by_weekend[[1]]
weekdays_only <- split_by_weekend[[2]]
```
